### 2013-07-19 Compiled/intro/board.tex
editorial board and reviewers

\documentclass[a4paper,16pt]{article} 
\usepackage{a4wide,times,latexsym}
\usepackage[compact,tiny]{titlesec}
\usepackage{url}
\usepackage{times}
\usepackage{graphics}
\usepackage{color}
%\usepackage[pdftex]{hyperref}
\usepackage{multirow}

\pagestyle{empty}
\begin{document}

\begin{center}
{\Huge \textbf{Editorial Board}}

\Large 
\vspace{1in}
\begin{tabular}{ll}
Ernst Biersack &	Eurecom \\
Olivier Bonaventure &	Universit$\acute{e}$ catholique de Louvain \\
Jon Crowcroft	&  University of Cambridge \\
Walid Dabbous	  & INRIA Sophia Antipolis \\
Bruce Davie	& VMware \\
Anja Feldmann	& Technische Universit$\ddot{a}$t Berlin\\
Timothy Griffin	& University of Cambridge \\
Hamed Haddadi &	Queen Mary University of London\\
Ramesh Johari	 & Stanford University\\
Srinivasan Keshav &	University of Waterloo\\
Jean-Yves Le Boudec	& $\acute{E}$cole polytechnique f$\acute{e}$d$\acute{e}$rale de Lausanne\\
Jennifer Rexford &	Princeton University\\
David Wetherall	& University of Washington\\
Walter Willinger	& AT\&T Labs Research\\
\end{tabular}
\end{center}

\newpage

\begin{center}
{\Huge \textbf{Review Board}}

\Large 
\vspace{1in}

\begin{tabular}{ll}
Mark Allman &	International Computer Science Institute \\
Emmanuel Baccelli    &	INRIA Saclay\\
Saleem Bhatti	& University of St Andrews \\
Olivier Bonaventure &	Universit$\acute{e}$ catholique de Louvain \\
Gonzalo Camarillo  &  Ericsson Research   \\
Dah Ming Chiu   &   The Chinese University of Hong Kong \\
Costas Courcoubetis  & Athens University of Economics and Business \\
Mark Crovella & Boston University \\
Jon Crowcroft	&  University of Cambridge \\
Bruce Davie	& VMware \\
Benoit Donnet  &   Universit$\acute{e}$ de Li$\grave{e}$ge  \\
Damien Fay & Bournemouth University  \\
Paul Francis & Max Planck Institute for Software Systems \\
Hamed Haddadi &	Queen Mary University of London\\
Luigi Iannone &  TELECOM ParisTech \\
Srinivasan Keshav    &	University of Waterloo\\
Olaf Maennel  & Loughborough University \\
Konstantinos Nikitopoulos & University College London \\
Costin Raiciu &  Universitatea Politehnica Bucuresti  \\
Jennifer Rexford &	Princeton University\\
Miguel Rio & University College London \\ 
Matthew Roughan & University of Adelaide \\
Jean-Louis Rougier  &  TELECOM ParisTech \\
Damien Saucez  & INRIA Sophia Antipolis \\
Aman Shaikh  & AT\&T Labs Research \\
Steve Uhlig & Queen Mary University of London\\
\end{tabular}
\end{center}

\end{document}