### 2013-07-19 Compiled/intro/intro.tex
first stab

Professors often rely on textbooks to teach undergraduate and graduate networking courses. While there are many good introductory textbooks, there are very few books on advanced networking topics that could be suitable to graduate courses in networking. To fill this gap, SIGCOMM Education Committee has launched a community project to develop a high-quality, open-source, edited eBook on “Recent Advances in Networking”. This eBook will be distributed online via the SIGCOMM website.

This eBook is composed of nine chapters chosen after a highly selective review process by the editorial board. The selected chapters cover advanced networking topics and accompanying teaching material (slides and exercises). All the source code of the eBook and the teaching material are kept on a version controlled repository that will be accessible for the entire SIGCOMM community. We expect that releasing such high quality teaching material will be beneficial for a large number of students and professors. The teaching material will be updated on a regular basis to reflect new advances in our field. We will also be adding new chapters on more emerging topics in the future volumes.

We wish to thank all there authors for providing such high quality chapters. We are also grateful to the reviewers and the editorial board for spending many hours on each chapter to ensure a coherent level amongst all the chapters. We hope you enjoy reading this eBook and find it a useful resource.

Hamed Haddadi, Olivier Bonaventure

August 2013