### 2013-07-22 Compiled/chapter8/future.tex

\clearpage
\section{Future Directions}
\label{sec:future}

There are some interesting tasks left for traffic matrix
research. The various algorithms and techniques described
here could be improved, though in many cases the improvements may be
relatively incremental given the success of existing approaches. More
interest may be found in extending the ideas and techniques used here
to new domains, and to evolving Internet traffic.

There are a few obvious cases (and no doubt many less obvious cases
that we have not thought of), for instance: multicast traffic has not,
to our knowledge, been studied in this way. Multicast is interesting
because it violates the {\em traffic conservation} assumption that
lies underneath many techniques for estimation and modelling of
traffic matrices. We could imagine modelling it by considering the
``flow'' to be the traffic on a multicast group, from say, one source,
to a set of destinations, and then stacking a vector with these. The
routing matrices now include elements for every link used (no longer
following a single path). The traffic ``matrix'' could then be a
column vector of the traffic on each of these flows. So the idea of
multicast traffic can fit into the structure we have talked about
here, but appropriate models for performing tasks such as inference
do not seem to exist.

It would also be very interesting to understand the way that CDNs are
affecting network traffic. A step in this direction, although not directly
on traffic, but more on the discovery of the content hosts is \cite{Ager11CDN}.
A CDN's typical goal is to bring content
closer to the user, thereby reducing network traffic. However, that
explicitly violates the ``friction free'' assumption in most gravity
models, and introduces distance as something to be modelled.

That leads naturally to the consideration of global traffic. Almost all
studies of traffic have concentrated on a single network no larger
than the national scale. That may still be very large -- for example,
several studies looked at Tier 1 providers in the USA, which for some
time dominated Internet traffic. However, although large, it was still
relatively homogeneous traffic between people speaking much the same
language(s) from place to place in the network. When we consider the
Internet globally, we may see that there are language or cultural
clusters where large groupings of traffic are focussed.

On a large scale, time zones also play a significant role. Traffic
patterns show strong cyclic behaviour based on user activity, but such
activity is strongly dependent on the local time zone. If traffic is
flowing from user to user, then this can result in strong apparent
locality effects, simply because people in the same time zone are more
likely to be awake at the same time \cite{gerber:03}.

While language and cultural focussing may be geographic in nature, it
might also be considered per network, and that leads to another topic
of some interest. Very few papers have tried to consider inter-AS
(also known as inter-domain) traffic in any detail. Exceptions are
Chang~\etal\cite{Chang05EmpiricalAS} (which presents suggestions for
estimating traffic based on models derived from business models and resulting
usage); Bharti~\etal\cite{Bharti10Invisible} (which considered
inference of hidden elements of this matrix using a subset of data),
Feldman~\etal\cite{Feldmann04Web} (which aimed to estimate a global
traffic matrix, but only in the limited domain of WWW traffic), and
Labovitz~\etal\cite{labovitz10:_inter_inter_domain_traff} (which
looked at inter-domain traffic from 110 network operators over a
two-year period, though not in the form of a matrix). Study of the
Internet's global traffic matrix is made difficult by the sheer scale
of the project: Labovitz~\etal studied 110 network operators over a
two-year period\footnote{To put this in context, there are tens of
  thousands of ASes in the Internet.} and to do so, collected over 200
Exabytes of data. Many network operators do not collect or store data
of the type required for such a study, and many more regard it as
proprietary or covered by privacy legislation with provisions such
that no researcher is ever likely to see it. So we can see that study
of the inter-domain matrix is likely to be a long-term, and rather
challenging project.

In addition, we know that the {\em traffic profile} (or mix of
applications) has changed fairly rapidly over time. It is likely this
trend will continue, and there are bound to be effects on traffic
patterns as a result. Peer-2-peer traffic significantly altered
traffic patterns when it appeared because it was more symmetric than
traditional (at the time) WWW traffic. However, in addition,
peer-2-peer applications have the potential to exploit locality
information to download from sources closer to the destination. This
could potentially change the ``no friction'' assumption in much the
same way that CDNs can, though in the early days it did not appear to
be the case \cite{gerber:03}. An example traffic matrix (drawn from
\cite{gerber:03}) showing normalised\footnote{The elements have been
  normalised by dividing each row by the row-sum, so that each element
  actually represents the probability that a packet enters the network
  at a given region $i$ will depart the network at region $j$.}
traffic between regions in a cable-network operator is given in
\autoref{tab:cable_tm}. The major deviation, in this data, from a pure
gravity model seemed to be time-zone differences.

\begin{table}[!h]
  \centering
  \begin{tabular}{r|rrrrrrrr}
    From/To & R1 & R2 & R3 & R4 & R5 & R6 & R7 & R8 \\
    \hline
    R1 & -     & 0.180 & 0.140 & 0.126 & 0.174 & 0.128 & 0.124 & 0.127 \\
    R2 & 0.172 & -     & 0.141 & 0.126 & 0.190 & 0.132 & 0.118 & 0.120 \\
    R3 & 0.132 & 0.120 & -     & 0.189 & 0.135 & 0.145 & 0.139 & 0.140 \\
    R4 & 0.107 & 0.111 & 0.182 & -     & 0.124 & 0.163 & 0.155 & 0.158 \\
    R5 & 0.161 & 0.180 & 0.136 & 0.132 & -     & 0.135 & 0.127 & 0.129 \\
    R6 & 0.107 & 0.108 & 0.145 & 0.155 & 0.125 & -     & 0.187 & 0.173 \\
    R7 & 0.107 & 0.106 & 0.137 & 0.157 & 0.127 & 0.182 & -     & 0.184 \\
    R8 & 0.109 & 0.111 & 0.127 & 0.161 & 0.128 & 0.178 & 0.185 & - \\
  \end{tabular}
  \caption{Normalised inter-regional traffic matrix from \cite{gerber:03}.}
  \label{tab:cable_tm}
\end{table}

% T = [
%  [ 0     , 0.180 , 0.140 , 0.126 , 0.174 , 0.128 , 0.124 , 0.127 ]
%  [ 0.172 , 0     , 0.141 , 0.126 , 0.190 , 0.132 , 0.118 , 0.120 ]
%  [ 0.132 , 0.120 , 0     , 0.189 , 0.135 , 0.145 , 0.139 , 0.140 ]
%  [ 0.107 , 0.111 , 0.182 , 0     , 0.124 , 0.163 , 0.155 , 0.158 ]
%  [ 0.161 , 0.180 , 0.136 , 0.132 , 0     , 0.135 , 0.127 , 0.129 ]
%  [ 0.107 , 0.108 , 0.145 , 0.155 , 0.125 , 0     , 0.187 , 0.173 ]
%  [ 0.107 , 0.106 , 0.137 , 0.157 , 0.127 , 0.182 , 0     , 0.184 ]
%  [ 0.109 , 0.111 , 0.127 , 0.161 , 0.128 , 0.178 , 0.185 , 0 ]
% ]

New traffic classes may change traffic matrices in the future, and
modelling these will be interesting. On the other side, applications
such as anomaly detection are likely to remain interesting due to
their immediate benefits to operators, but the most overlooked task is
traffic matrix synthesis. As mentioned in the previous section, the
lack of real world traffic matrix datasets motivates the use of
artificial traffic matrices to provide some degree of approximation in
network traffic planning, provisioning and engineering tasks.  It is
an important area to concentrate on, considering its usefulness to
network operators.

%Synthesis means generating artificial traffic matrices, typically for
%use in simulations. There are only, to our knowledge, two papers
%\cite{Nucci05TMSynth,Roughan05GravSynth} on synthesising Internet
%traffic matrices, but there are already quite a few where synthetic
%traffic matrices were used, and this demand for such matrices will
%continue.
%
%Synthesis is not demanding in some ways. Traffic matrices are usually
%relatively small (compared to other types of traffic data), when
%measured at a reasonable level of aggregation and time scale. However,
%in other ways these matrices are quite challenging. For instance:
%\begin{itemize}
%
%\item we have few sets of traffic matrix data, and even fewer that are
%  public, and somehow need to use these to estimate properties of
%  these complex, high-dimensional objects;
%
%\item there is a real relationship between topology and traffic
%  (although we would like a traffic matrix to be invariant to the
%  topology, there are clear cases where, particularly IE matrices are
%  not);
%
%\item traffic matrices come in a wide variety of types (at different
%  levels of aggregation, for particular applications and so on) and it
%  is unlikely that one model fits all; and 
%
%\item there are a number of conflicting goals in synthesis, \eg~to
%  generate variability, but well ``matched'' to real traffic
%  matrices. 
%
%\end{itemize}
%However, there is considerable hope that progress can be made in terms
%of generating synthetic matrices, both for green-fields network design
%\cite{Kowalski95TeleModel}, and for simulation in general. 
%
%The major use of synthesis is in simulation. In many cases a traffic
%matrix is enough for a simulation, but in others, we need to translate
%this into packets (or at least connections). The analogue in
%transportation modelling is often called a {\em micro-simulation}
%model. Here, the problem becomes one of taking a {\em demand} matrix
%(remember, most of the work here is related to traffic, not demand), and
%translating this into carried load. We know how to do that (using
%simulation tools such as \verb|ns|) but doing it efficiently is
%difficult. One paper \cite{sommers11:_effic} starts to tackle this
%problem, but as in the work of transportation modelling, there is
%considerable scope for advanced scalable micro-simulation of traffic.
%
%Another use for synthetic traffic matrices is in the further task of
%synthetic topology generation, but we shall leave discussion of this
%topic to another section of this book.