### 2013-07-22 originals/chapter1/ebook.tex

% This work is licensed under
% http://creativecommon.org/licenses/by/3.0/
\documentclass{article}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{color}
\usepackage{url}
\hyphenation{wrap-around}

%this is a nice-looking bullet list:
%\begin{list}{$\bullet$}{
%                      \setlength{\topsep}{0in}
%                      \setlength{\itemsep}{0in}
%                      \setlength{\parsep}{0in}
%                      \setlength{\leftmargin}{.13in}
%                     }
%\item
%\end{list}

%this is a two-column figure:
%(for a one-column figure, remove *)
%figure must be in ???.pdf
%\begin{figure*}[hbt]
%\begin{center}
%\includegraphics[scale=1.00]{???}
%\end{center}
%\caption{???}
%\label{fig:???}
%\end{figure*}

%this is a two-column, floating table:
%(for a one-column table, remove *)
%(for a non-floating table, remove table environment)
%\begin{table*}
%\begin{tabular}
%\end{tabular}
%\end{table*}

%JEN: for new text added by jen.  can simply remove \textcolor{red} to
%get rid of red color, or remove the \jen{} where it appears in the text.
\newcommand{\jen}[1]{\textcolor{red}{#1}}

\title{The Design Space of Network Mobility}

\author{
Pamela Zave \\
AT\&T Laboratories---Research \\
Florham Park, New Jersey, USA \\
{\tt pamela@research.att.com}
\and
Jennifer Rexford \\
Princeton University \\
Princeton, New Jersey, USA \\
{\tt jrex@cs.princeton.edu}
}

\begin{document}
\thispagestyle{empty}
\maketitle

\begin{abstract}
While the Internet is increasingly mobile, seamless mobility is
difficult to implement at Internet scale.  Over the years, standards
bodies and the research community have introduced a large and
confusing collection of mobility proposals that are difficult to
compare.  In this tutorial, we present these mobility proposals
in a uniform framework, called the geomorphic view of networking.
The geomorphic view shows that there are two distinct patterns
for implementing mobility, each with its own range of 
design choices and cost-benefit trade-offs.
We use these patterns to classify and explain
a representative sample of mobility
mechanisms, abstractly yet precisely.
The patterns also serve as a basis for evaluating properties of
these mechanisms such as resource costs and scalability,
and for considering composition of mobility mechanisms.
\end{abstract}

\input{ebook-sec1}

\input{ebook-sec2}

\input{ebook-sec3}

\input{ebook-sec4}

\input{ebook-sec5}

\input{ebook-sec6}

\input{ebook-sec7}

\input{ebook-sec8}

\input{ebook-sec9}

\bibliographystyle{IEEEtran}
\bibliography{ebook}

\end{document}