### 2013-07-22 Compiled/chapter4/sections/Introduction.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Recent traffic studies~\cite{TrafficTypesGrowth:2011,arbor,PADIS2010} show that
a large fraction of Internet traffic is due to content delivery and is
originated by a small number of Content Delivery Infrastructures (CDIs). Major
CDIs include highly popular rich-media sites like YouTube and Netflix,
One-Click Hosters (OCHs), \eg RapidShare~\cite{OCH-performance},
Content Delivery Networks (CDNs) such as Akamai and Limelight, and
hyper-giants, \eg Google, Yahoo!, and Microsoft.  Gerber and Doverspike~\cite
{TrafficTypesGrowth:2011} report that a few CDIs account for more than half of
the traffic of a US-based Tier-1 carrier.  Poese \etal~\cite{PADIS2010} report
a similar observation from the traffic of a European Tier-1 carrier. Labovitz
\etal~\cite{arbor} infer that more than 10\% of the total Internet inter-domain
traffic originates from Google, and Akamai claims to deliver more than 20\% of
the total Internet Web traffic~\cite{Akamai-Network}. Netflix alone, a company
that offers a high definition video video-on-demand streaming service, is
responsible for a significant fraction of the traffic in North America ISPs
during peak hours~\cite{sandvine,Netflix-CDN-presentation}.

To cope with the increasing demand for content, CDIs have deployed massively
distributed server infrastructures to replicate content and make it accessible
from different locations on the Internet \cite{CDNsec2009}. These
infrastructures have multiple choices as to how and where to place their
servers. As described in~\cite {ImprovingPerformanceInternet2009}, the main
approaches are (1) centralized hosting, (2) data center-based CDIs, (3)
edge-cache-based CDIs, and (4) peer-to-peer (P2P) networks. Approaches~2 and~3
scale content delivery by distributing the content onto dedicated
infrastructures.  These infrastructures can be composed of a few large data
centers, a large number of edge caches, or any combination thereof.

To complicate matters further, some of these infrastructures are entangled with
the very infrastructures that provide network connectivity to end-users. For
example, one of the largest players in content delivery, Akamai, operates more
than 120,000 servers in more than 2,000 locations across nearly 1,150 ISP
networks~\cite {Akamai-Network,Akamai-website}. Google is reported to operate
tens of data centers and front-end server clusters
worldwide~\cite{MovingBeyondE2E2009,Tariq:What-if,Google-Datacenters}.
Microsoft has deployed its content delivery infrastructure in 24 locations
around the world~\cite{azure}. Amazon maintains at least 5 large data centers
and caches in at least 21 locations around the world~\cite{amazon}. Limelight
operates thousands of servers in more than 22 delivery centers and connects
directly to 600 networks worldwide~\cite{LLNetworks}. Last but not least, P2P
networks rely on a huge number of end users to store, replicate, and distribute
content.

Despite the significant entanglement between the infrastructures that deliver
content and the network connectivity fabric, our knowledge of their
interactions is largely through the literature on network interconnections, \eg
see the recent book by W. Norton~\cite{peering-playbook}. Given the nature of
network interconnections, previous work has studied the interactions from an
economic perspective~\cite
{internet-economics,interactions-economics,content-oriented-economics}. The
limited knowledge available about the settlements between networks have led
researchers to try to reason about why peering choices are made~\cite
{ToPeerOrNot:Infocom2006} and what drives the evolution of the
Internet~\cite{evolution}.

Most of the literature has considered the interactions between content and the
network indirectly, \eg through peerings and traffic measurements, despite
recent changes in Internet traffic~\cite{TrafficTypesGrowth:2011,arbor} that
have shown the importance of content and applications.  The observed changes in
traffic, either through direct traffic measurements~\cite
{TrafficDemand:ToN2001,Feldmann:Kammenhuber:2004,uhlig-public,arbor,IXPSIGCOMM2012},
or through
inference~\cite{medina_tm:2002,zhang_tm:2003,zhang_link:2003,gunnar_estimation:2004,dina_distributed:2004,soule_tm:2005}
have repeatedly shown how volatile traffic can be. With the rise of
user-generated content and large shifts in content popularity, traffic
volatility has become especially relevant.

Handling changes in traffic has traditionally been done through traffic
engineering (TE). Initially, traffic engineering was used by large network
operators to optimize the utilization of their networks
\cite{Awduche_OverviewTE:2002}. The vast majority of the traffic engineering
literature has therefore focused on traffic engineering inside a single network
\cite{netscope,FT00,Xiao_Traffic:2000,Aukia_RATES:2000,MPLS:Infocom2000,FT01}.
In reality, most of the traffic in the Internet is exchanged between different
networks \cite{arbor}, and especially directly between data centers and
residential ISPs \cite{IXPSIGCOMM2012}. Organizations that originate a lot of
content, \eg Google, connect directly to a large number of other networks
\cite{arbor}, and need to optimize how content leaves their networks.
Organizations that provide Internet access to broadband or mobile users
typically wish to optimize how traffic enters their networks, as most users
still download more content than they upload. In between, transit ISPs try to
balance the load of the traffic exchanged between the networks they connect.

Traditional traffic engineering aims at reducing the likelihood that
bottlenecks arise inside a given network due to mismatches between network
provisioning and expected demand. Changes in network provisioning are slow,
taking place over time scales of weeks or months. Popular content, on the other
hand, generates bursts in demand over much smaller time scales, \eg hours or
minutes. Today's Internet requires much more reactive network control
techniques than those we have today, and these techniques must take content
delivery into consideration. A few steps have been made in this direction.
Indeed, collaborative approaches
\cite{TECDN,CooperativeSettlement:ToN,Cate-CCR} have been proposed to help deal
with the traffic generated by content delivery infrastructures. Even in the
case of P2P, portals have been proposed to allow P2P applications and users to
communicate with ISPs to receive updated views of their networks~\cite{p4p}. In
broad terms, all information CDIs are missing today for optimizing their
operations is available to ISPs. Combined with the already proposed schemes for
collaboration, it is surprising how little real collaboration is performed in
today's Internet between these parties.

In this chapter, we analyze the operation of CDIs as well as network operators.
The analysis demonstrates the potential for fruitful collaboration.  We argue
that for collaboration to become more common, it is important for every party
in the content delivery landscape, \ie the content delivery infrastructures,
the network operators, and the end users, to benefit.  Finally, we present, in
depth, two systems that have incentives for every party and that can readily be
used today.