### 2013-07-23 Compiled/intro/chapterlist.tex
Full compiled

%\documentclass[a4paper,16pt]{article} 
%\usepackage{a4wide,times,latexsym}
%\usepackage[compact,tiny]{titlesec}
%\usepackage{url}
%\usepackage{times}
%\usepackage{graphics}
%\usepackage{color}
%%\usepackage[pdftex]{hyperref}
%\usepackage{multirow}

%\pagestyle{empty}
%\begin{document}

\begin{center}
{\Huge \textbf{List of Chapters}}
\end{center}

\vspace{1cm}

\large

\begin{enumerate}
\item \textbf{Internet Topology Research Redux}\\
Walter Willinger, Matthew Roughan

\item \textbf{Recent Advances in Transport Protocols}\\
Olivier Bonaventure, Janardhan Iyengar, Costin Raiciu

\item \textbf{Internet Traffic Matrices: A Primer} \\
Paul Tune, Matthew Roughan

\item \textbf{Optimizing and Modeling Dynamics in Networks } \\
Ibrahim Matta

\item \textbf{Smart Data Pricing (SDP): Economic Solutions to Network Congestion} \\
Soumya Sen, Carlee Joe-Wong, Sangtae Ha, Mung Chiang

\item \textbf{MPLS Virtual Private Networks } \\
Luca Cittadini, Giuseppe Di Battista, Maurizio Patrignani

\item \textbf{Collaboration Opportunities for Content Delivery and Network \\ Infrastructures} \\
Benjamin Frank, Ingmar Poese, Georgios Smaragdakis, Anja Feldmann, Bruce M. Maggs, Steve Uhlig, Vinay Aggarwal, and Fabian Schneider

\item \textbf{The Design Space of Network Mobility} \\
Pamela Zave, Jennifer Rexford

\item \textbf{Enabling Multihop Communication in Spontaneous Wireless Networks} \\
Juan Antonio Cordero, Jiazi Yi, Thomas Clausen, Emmanuel Baccelli

\end{enumerate}

%\end{document}