### 2013-07-23 Compiled/ebook.tex
top-level ebook

\documentclass{book}

%styles

\usepackage[final]{pdfpages}

% from chapter 1

%\usepackage{graphicx}
%\usepackage{latexsym}
%\usepackage{amsmath}
%\usepackage{color}
%\usepackage{url}
%\hyphenation{wrap-around}

\title{Recent Advances in Networking}

\begin{document}

\maketitle

%\frontmatter

\input{preface/preface.tex}

\newpage

\input{intro/intro.tex}
\newpage

\input{intro/board.tex}
\newpage

\input{intro/chapterlist.tex}

\newpage

%\mainmatter

\includepdf[pages=-]{chapter1/chapter1.pdf}
\includepdf[pages=-]{chapter2/chapter2.pdf}
\includepdf[pages=-]{chapter3/chapter3.pdf}
\includepdf[pages=-]{chapter4/chapter4.pdf}
\includepdf[pages=-]{chapter5/chapter5.pdf}
\includepdf[pages=-]{chapter6/chapter6.pdf}
\includepdf[pages=-]{chapter7/chapter7.pdf}
\includepdf[pages=-]{chapter8/chapter8.pdf}
\includepdf[pages=-]{chapter9/chapter9.pdf}

%\backmatter

\end{document}