### 2013-07-24 Compiled/chapter2/copyright.tex
added page numbers and initial support for copyright, urls are now clickable

\usepackage[absolute,showboxes]{textpos}

%set unit to be pagewidth and height, and increase inner margin of box
\setlength{\TPHorizModule}{\paperwidth}
\setlength{\TPVertModule}{\paperheight}
\TPMargin{5pt}

%define \copyrightstatement command for easier use
\newcommand{\copyrightstatement}{
    \begin{textblock}{0.7}(0.08,0.9)    % tweak here: {box width}(leftposition, rightposition)
         \noindent
         \footnotesize
         P. Zave, J. Rexford, ``The Design Space of Network
         Mobility'', in H. Haddadi, O. Bonaventure (Eds.), \emph{Recent
         Advances in Networking}, (2013), pp. 5-45. Licensed under a
         creative commons license and available from \url{http://www.sigcomm.org}
    \end{textblock}
}