### 2013-07-24 Compiled/commonstyle.tex

\usepackage{fullpage} % use entire page layout
\usepackage{times}     % smaller and better fonts that the standard latex ones
\usepackage[letterpaper]{geometry} % force letter format
\usepackage[colorlinks=true,linkcolor=black,urlcolor=blue,anchorcolor=black,citecolor=black]{hyperref}