### 2013-07-24 Compiled/ebook.tex

\documentclass{book}

%styles
\input{commonstyle}

%\hypersetup{pdftitle={Recent Advances in Networking},\
%           pdfauthor={Hamed Haddadi and Olivier Bonaventure (Editors)}}

\usepackage[final]{pdfpages}

%\includepdfset{pagecommand=\thispagestyle{plain}}

% from chapter 1

%\usepackage{graphicx}
%\usepackage{latexsym}
%\usepackage{amsmath}
%\usepackage{color}
%\usepackage{url}
%\hyphenation{wrap-around}

\title{Recent Advances in Networking}

\author{
Hamed Haddadi,  Olivier Bonaventure (Editors)\\
}

\begin{document}

\date{August 2013}

\maketitle
\pagenumbering{roman}

%\frontmatter

\input{intro/intro.tex}
\newpage

\input{preface/preface.tex}

\newpage

\input{intro/board.tex}
\newpage

\input{intro/chapterlist.tex}

\newpage

\pagenumbering{arabic}

%\mainmatter

\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter1/chapter1.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter2/chapter2.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter3/chapter3.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter4/chapter4.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter5/chapter5.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter6/chapter6.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter7/chapter7.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter8/chapter8.pdf}
\includepdf[pages=-,pagecommand=\thispagestyle{plain}]{chapter9/chapter9.pdf}

%\backmatter

\end{document}