### 2013-07-24 Compiled/commonstyle.tex
new bibliography style with hyperlinks, based on http://www.math.tamu.edu/~comech/tools/bibtex-doi-eprint/

\usepackage{fullpage} % use entire page layout
\usepackage{times}     % smaller and better fonts that the standard latex ones
\usepackage[letterpaper]{geometry} % force letter format
\usepackage[colorlinks=true,linkcolor=black,urlcolor=blue,anchorcolor=black,citecolor=black,backref=none]{hyperref}
\usepackage[usenames,dvipsnames]{color}
%% Disable backref and make \href colors more decent:
\definecolor{MyDarkBlue}{rgb}{0,0.1,0.7}
\hypersetup{pdfborder={0 0 0},colorlinks,breaklinks=true,
  urlcolor={MyDarkBlue},citecolor={MyDarkBlue},linkcolor={MyDarkBlue} }