### 2013-07-25 Compiled/chapter5/main.tex
fixes

%\newcommand{\CLASSINPUTinnersidemargin}{0.7in}
%\newcommand{\CLASSINPUToutersidemargin}{0.7in}

\documentclass{article}

\input{../commonstyle}

\input{copyright}

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{array}
\usepackage{fancyvrb}
\usepackage{boxedminipage}
\usepackage{color}
\usepackage{paralist}
\usepackage{algorithmic}
\usepackage{slashbox}
\usepackage{xspace}
\usepackage{url}
\usepackage{subfig}
\usepackage{amsthm}
\newcommand{\ed}[1]{\textsf{\textcolor{red}{[#1]}}}
\newcommand{\remove}[1]{}
\usepackage{epstopdf}
\epstopdfsetup{update}

\title{Recent Advances in Reliable Transport Protocols\footnote{
Parts of the text in this chapter have appeared in the following
 publications by the same author(s): \cite{login}, \cite{mptcp-cc},
 \cite{raiciu2012hard}, ~\cite{nowlan12fitting} 
and~\cite{iyengar10minion}.}
}

\author{Costin Raiciu, Janardhan Iyengar, Olivier Bonaventure}

\date{}

\begin{document}

\pagestyle{empty}

\maketitle

\thispagestyle{empty}

\begin{abstract}
Transport protocols play a critical role in today's Internet. This
chapter first looks at the recent of the reliable transport
protocols. It then explains the growing impact of middleboxes on the
evolvability of these protocols. 
Two recent protocol extensions,
Multipath TCP and Minion,
which were both designed to extend the current Transport Layer 
in the Internet
are then described.
\end{abstract}

\section{Introduction}
%%%%%%%%%%%%%%%%%%

\input{intro.tex}

\section{Today's Internet}\label{section:today}
%%%%%%%%%%%%%%%%%%%%%

\input{internet.tex}

%\section{Structuring the transport layer}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{structure.tex}

\section{Multipath TCP}\label{section:mptcp}

%%%%%%%%%%%%%%%%%%%%

\input{mptcp.tex}

\section{Minion}\label{section:minion}
%%%%%%%%%%%%%

\input{minion.tex}

%\section{A glimpse into future transport protocols}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{future.tex}

\section{Conclusion}\label{section:conclusion}
%%%%%%%%%%%%%%%%%

\input{conclusion.tex}

%\bibliographystyle{plain}

\input{../commonbibstyle}
\bibliography{ebook,rfc}

\appendix

\section{Exercises}
%%%%%%%%%%%%%%%%

\input{exercises.tex}

\end{document}