### 2013-07-26 Compiled/chapter1/copyright.tex
copyright boxes in all chapters

\usepackage[absolute,showboxes]{textpos}

%set unit to be pagewidth and height, and increase inner margin of box
\setlength{\TPHorizModule}{\paperwidth}
\setlength{\TPVertModule}{\paperheight}
\TPMargin{5pt}

%define \copyrightstatement command for easier use
\newcommand{\copyrightstatement}{
    \begin{textblock}{0.65}(0.17,0.9)    % tweak here: {box width}(hposition, vposition)
         \noindent
         \footnotesize
         P. Zave, J. Rexford, ``The Design Space of Network
         Mobility'', in H. Haddadi, O. Bonaventure (Eds.), \href{\ebookurl}{\emph{Recent
         Advances in Networking}}, (2013), pp. xx-yy. Licensed under a
         \href{http://creativecommons.org/licenses/by-sa/3.0/}{CC-BY-SA}
         Creative Commons license. 
    \end{textblock}
}